\pkg{cubble}: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data

### 2022-06-08 2-0-cube.Rmd
tweakinf first few sections

The most common data cube using the three axes (time, latitude, longitude) could be considered as taking snapshots of the space and stacking them across time. Here we define the spatio-temporal cube slightly differently by considering the three axes to be, time, site and  variable/s. That is illustrated in the leftmost cube in Figure \ref{fig:cube-def}. The time axis is the same in both versions, while the site axis now captures both latitude and longitude. Finally, variables are stacked on this space-time canvas, with one observation per site and time point. This notion is adopted to avoid using hypercubes when describing multivariate spatio-temporal data and is the conceptual framework behind the \pkg{cubble} objects. With this conceptual model, operations on spatio-temporal data can be mapped to operations on the cube and the rest of Figure \ref{fig:cube-def} show examples of slicing on site, time, and variable.

While the data cube model is conceptually convenient for spatio-temporal data, a 3D data array is not sufficiently rich for data wrangling, for several reasons. Although arrays can be efficient for the computation on numerical values, spatio-temporal data typically includes various types of variables. For example, character strings and specific datetime classes are common. In addition, it will be generally useful to be able to create new variables which is trickier to manage in an array. Thus for convenient wrangling, we have opted for a list object as the `cubble` structure.

>>>>

The most common space-time cube uses the three axes, time, latitude, longitude, and can be considered stacking space across time. Ours is a multivariate spatio-temporal cube with the three axes defined to be time, site and variables, as illustrated in the leftmost column of Figure \ref{fig:cube-def}. The time axis is the same in both versions, while the site axis now captures both latitude and longitude. Finally, variables are stacked on this space-time canvas, with one observation per site and time point. This notion is adopted to avoid using hypercubes when describing multivariate spatio-temporal data and is the conceptual framework behind the \pkg{cubble} objects. With this conceptual model, operations on spatio-temporal data can be mapped to operations on the cube and the rest of Figure \ref{fig:cube-def} show examples of slicing on site, time, and variable.

While the data cube model is conceptually convenient for spatio-temporal data, a 3D data array is not sufficiently rich for data wrangling, for several reasons. Although arrays can be efficient for the computation on numerical values, spatio-temporal data typically includes various types of variables. For example, character strings and specific datetime classes are common. In addition, it will be generally useful to be able to create new variables which is trickier to manage in an array. Thus for convenient wrangling, we have opted to create a special `cubble` class.

### 2022-06-08 2-the-cubble-pkg.Rmd
tweakinf first few sections

A cubble object is an S3 object [@advr] built on top of the \code{tibble} class to organise spatio-temporal data. Since spatial and temporal variables are usually recorded in different observational units, the \code{cubble} class uses an attribute, **form**,  to arrange the data into a spatial or temporal tidy way. The **form** attributes can take a value of either "nested" or "long". The nested cubble is a subclass of \code{rowwise_df}. It arranges each spatial site in a row and uses list columns to store feature geometry and temporal information. The long cubble, instead, is a subclass of \code{grouped_df}, which expands the temporal information into the long form and stores the spatial information in a unique attribute to the long cubble, **spatial**.

@@

The main functions in the cubble package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()}. The following sections explain their role, why the new cubble structure is needed and how the package relates to existing packages for spatial and temporal data analysis.

>>>>

A \code{cubble} object is an S3 class [@advr] built on the \code{tibble} class, specifically to organise spatio-temporal data. The \code{cubble} class uses an attribute "form",  to arrange the spatial or temporal data components tidily. The form attribute can take a value of either "nested" or "long". The nested cubble is a subclass of \code{rowwise_df}. It arranges each spatial site in a row, and uses list columns to store the feature geometry and the temporal information. The long cubble is a subclass of \code{grouped_df}, which expands the temporal information into the long form and stores the spatial information in a "spatial" attribute.

@@

The main functions in the package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()}. The following sections explain their role, why the new cubble structure is needed and how the package relates to existing packages for spatial and temporal data analysis.

### 2022-06-08 jss/cubble-jss.tex
tweakinf first few sections

The most common data cube using the three axes (time, latitude, longitude) could be considered as taking snapshots of the space and stacking them across time. Here we define the spatio-temporal cube slightly differently by considering the three axes to be, time, site and variable/s. That is illustrated in the leftmost cube in Figure \ref{fig:cube-def}. The time axis is the same in both versions, while the site axis now captures both latitude and longitude. Finally, variables are stacked on this space-time canvas, with one observation per site and time point. This notion is adopted to avoid using hypercubes when describing multivariate spatio-temporal data and is the conceptual framework behind the \pkg{cubble} objects. With this conceptual model, operations on spatio-temporal data can be mapped to operations on the cube and the rest of Figure \ref{fig:cube-def} show examples of slicing on site, time, and variable.

While the data cube model is conceptually convenient for spatio-temporal data, a 3D data array is not sufficiently rich for data wrangling, for several reasons. Although arrays can be efficient for the computation on numerical values, spatio-temporal data typically includes various types of variables. For example, character strings and specific datetime classes are common. In addition, it will be generally useful to be able to create new variables which is trickier to manage in an array. Thus for convenient wrangling, we have opted for a list object as the \texttt{cubble} structure.

@@

A cubble object is an S3 object \citep{advr} built on top of the \code{tibble} class to organise spatio-temporal data. Since spatial and temporal variables are usually recorded in different observational units, the \code{cubble} class uses an attribute, \textbf{form}, to arrange the data into a spatial or temporal tidy way. The \textbf{form} attributes can take a value of either ``nested'' or ``long''. The nested cubble is a subclass of \code{rowwise_df}. It arranges each spatial site in a row and uses list columns to store feature geometry and temporal information. The long cubble, instead, is a subclass of \code{grouped_df}, which expands the temporal information into the long form and stores the spatial information in a unique attribute to the long cubble, \textbf{spatial}.

The main functions in the cubble package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()}. The following sections explain their role, why the new cubble structure is needed and how the package relates to existing packages for spatial and temporal data analysis.

@@

\begin{CodeChunk}
\begin{CodeOutput}
# A tibble: 1,830 x 10
  id         lat  long  elev name  wmo_id date        prcp  tmax  tmin
  <chr>    <dbl> <dbl> <dbl> <chr>  <dbl> <date>     <dbl> <dbl> <dbl>
1 ASN0000~ -31.9  116.  15.4 pert~  94610 2020-01-01     0  31.9  15.3
2 ASN0000~ -31.9  116.  15.4 pert~  94610 2020-01-02     0  24.9  16.4
3 ASN0000~ -31.9  116.  15.4 pert~  94610 2020-01-03     6  23.2  13  
4 ASN0000~ -31.9  116.  15.4 pert~  94610 2020-01-04     0  28.4  12.4
5 ASN0000~ -31.9  116.  15.4 pert~  94610 2020-01-05     0  35.3  11.6
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_nested <- climate_flat |>
+   as_cubble(key = id, index = date, coords = c(long, lat))
R> cubble_nested
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts      
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>  
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble>
2 ASN00010311 -31.9  117. 179   york            94623 <tibble>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble>
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_nested |>
+   mutate(rain_day = sum(ts$prcp != 0))
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name          wmo_id ts       rain_day
  <chr>       <dbl> <dbl> <dbl> <chr>          <dbl> <list>      <int>
1 ASN00009021 -31.9  116.  15.4 perth airport  94610 <tibble>      104
2 ASN00010311 -31.9  117. 179   york           94623 <tibble>       89
3 ASN00010614 -32.9  117. 338   narrogin       94627 <tibble>       90
4 ASN00014015 -12.4  131.  30.4 darwin airpo~  94120 <tibble>      106
5 ASN00015131 -17.6  134. 220   elliott        94236 <tibble>       63
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_long <- cubble_nested |> face_temporal()
R> cubble_long
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id
#   [dbl]
  id          date        prcp  tmax  tmin
  <chr>       <date>     <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3
2 ASN00009021 2020-01-02     0  24.9  16.4
3 ASN00009021 2020-01-03     6  23.2  13  
4 ASN00009021 2020-01-04     0  28.4  12.4
5 ASN00009021 2020-01-05     0  35.3  11.6
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_back <- cubble_long |> face_spatial()
R> cubble_back
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts      
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>  
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble>
2 ASN00010311 -31.9  117. 179   york            94623 <tibble>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble>
\end{CodeOutput}
\begin{CodeInput}
R> identical(cubble_nested, cubble_back)
\end{CodeInput}
\begin{CodeOutput}
[1] TRUE
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_unfold <- cubble_long |> unfold(long, lat)
R> cubble_unfold
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id
#   [dbl]
  id          date        prcp  tmax  tmin  long   lat
  <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3  116. -31.9
2 ASN00009021 2020-01-02     0  24.9  16.4  116. -31.9
3 ASN00009021 2020-01-03     6  23.2  13    116. -31.9
4 ASN00009021 2020-01-04     0  28.4  12.4  116. -31.9
5 ASN00009021 2020-01-05     0  35.3  11.6  116. -31.9
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_unfold |> face_spatial() |> face_temporal()
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id
#   [dbl]
  id          date        prcp  tmax  tmin
  <chr>       <date>     <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3
2 ASN00009021 2020-01-02     0  24.9  16.4
3 ASN00009021 2020-01-03     6  23.2  13  
4 ASN00009021 2020-01-04     0  28.4  12.4
5 ASN00009021 2020-01-05     0  35.3  11.6
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # example with a tsibble created from climate_flat
R> raw <- climate_flat |> tsibble::as_tsibble(key = id, index = date)
R> dt <-  raw |> cubble::as_cubble(coords = c(long, lat))
R> dt
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts      
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>  
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tbl_ts>
2 ASN00010311 -31.9  117. 179   york            94623 <tbl_ts>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tbl_ts>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tbl_ts>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tbl_ts>
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> covid |> head(5)
\end{CodeInput}
\begin{CodeOutput}
# A tsibble: 5 x 5 [1D]
# Key:       lga [1]
# Groups:    lga, source [1]
  date       lga        source                            n roll_mean
  <date>     <chr>      <chr>                         <int>     <dbl>
1 2022-01-01 Alpine (S) Contact with a confirmed case     1        NA
2 2022-01-02 Alpine (S) Contact with a confirmed case     2        NA
3 2022-01-03 Alpine (S) Contact with a confirmed case     4        NA
4 2022-01-04 Alpine (S) Contact with a confirmed case     4        NA
5 2022-01-05 Alpine (S) Contact with a confirmed case     2        NA
\end{CodeOutput}
\begin{CodeInput}
R> lga |> head(5)
\end{CodeInput}
\begin{CodeOutput}
Simple feature collection with 5 features and 7 fields
Geometry type: MULTIPOLYGON
Dimension:     XY
Bounding box:  xmin: 142.3535 ymin: -38.67876 xmax: 147.3909 ymax: -36.39269
Geodetic CRS:  WGS 84
    lga_code_2018            lga state_code_2016 state_name_2016
132         20110     Alpine (S)               2        Victoria
133         20260    Ararat (RC)               2        Victoria
134         20570   Ballarat (C)               2        Victoria
135         20660    Banyule (C)               2        Victoria
136         20740 Bass Coast (S)               2        Victoria
    areasqkm_2018 cent_long  cent_lat                       geometry
132     4788.1568  146.9742 -36.85357 MULTIPOLYGON (((146.7258 -3...
133     4211.1171  142.8432 -37.47271 MULTIPOLYGON (((143.1807 -3...
134      739.0321  143.7815 -37.49286 MULTIPOLYGON (((143.6622 -3...
135       62.5402  145.0851 -37.73043 MULTIPOLYGON (((145.1357 -3...
136      865.8095  145.5581 -38.50730 MULTIPOLYGON (((145.5207 -3...
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cb <- as_cubble(
+   list(spatial = lga, temporal = covid),
+   key = lga, index = date, coords = c(cent_long, cent_lat)
+   )
\end{CodeInput}
\begin{CodeOutput}
! Some sites in the temporal table don't have spatial information
\end{CodeOutput}
\begin{CodeOutput}
! Some sites in the spatial table don't have temporal information
\end{CodeOutput}
\begin{CodeOutput}
! Use argument `output = "unmatch"` to check on the unmatched key
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeOutput}
# cubble:   id [8]: nested form
# bbox:     [144.52, -37.73, 146.06, -36.55]
# temporal: date [date], matched_var [dbl]
  id          name        lat  long type   dist group ts       n_match
  <chr>       <chr>     <dbl> <dbl> <chr> <dbl> <int> <list>     <int>
1 405234      SEVEN CR~ -36.9  146. river  6.15     5 <tibble>      21
2 404207      HOLLAND ~ -36.6  146. river  8.54    10 <tibble>      21
3 ASN00082042 strathbo~ -36.8  146. clim~  6.15     5 <tibble>      21
4 ASN00082170 benalla ~ -36.6  146. clim~  8.54    10 <tibble>      21
5 230200      MARIBYRN~ -37.7  145. river  6.17     6 <tibble>      19
6 ASN00086038 essendon~ -37.7  145. clim~  6.17     6 <tibble>      19
7 406213      CAMPASPE~ -37.0  145. river  1.84     1 <tibble>      18
8 ASN00088051 redesdale -37.0  145. clim~  1.84     1 <tibble>      18
\end{CodeOutput}
\end{CodeChunk}

>>>>

The most common space-time cube uses the three axes, time, latitude, longitude, and can be considered stacking space across time. Ours is a multivariate spatio-temporal cube with the three axes defined to be time, site and variables, as illustrated in the leftmost column of Figure \ref{fig:cube-def}. The time axis is the same in both versions, while the site axis now captures both latitude and longitude. Finally, variables are stacked on this space-time canvas, with one observation per site and time point. This notion is adopted to avoid using hypercubes when describing multivariate spatio-temporal data and is the conceptual framework behind the \pkg{cubble} objects. With this conceptual model, operations on spatio-temporal data can be mapped to operations on the cube and the rest of Figure \ref{fig:cube-def} show examples of slicing on site, time, and variable.

While the data cube model is conceptually convenient for spatio-temporal data, a 3D data array is not sufficiently rich for data wrangling, for several reasons. Although arrays can be efficient for the computation on numerical values, spatio-temporal data typically includes various types of variables. For example, character strings and specific datetime classes are common. In addition, it will be generally useful to be able to create new variables which is trickier to manage in an array. Thus for convenient wrangling, we have opted to create a special \texttt{cubble} class.

@@

A \code{cubble} object is an S3 class \citep{advr} built on the \code{tibble} class, specifically to organise spatio-temporal data. The \code{cubble} class uses an attribute ``form'', to arrange the spatial or temporal data components tidily. The form attribute can take a value of either ``nested'' or ``long''. The nested cubble is a subclass of \code{rowwise_df}. It arranges each spatial site in a row, and uses list columns to store the feature geometry and the temporal information. The long cubble is a subclass of \code{grouped_df}, which expands the temporal information into the long form and stores the spatial information in a ``spatial'' attribute.

The main functions in the package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()}. The following sections explain their role, why the new cubble structure is needed and how the package relates to existing packages for spatial and temporal data analysis.

@@

\begin{CodeChunk}
\begin{CodeOutput}
# A tibble: 1,830 x 10
  id            lat  long  elev name         wmo_id date        prcp  tmax  tmin
  <chr>       <dbl> <dbl> <dbl> <chr>         <dbl> <date>     <dbl> <dbl> <dbl>
1 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-01     0  31.9  15.3
2 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-02     0  24.9  16.4
3 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-03     6  23.2  13  
4 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-04     0  28.4  12.4
5 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 2020-01-05     0  35.3  11.6
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_nested <- climate_flat |>
+   as_cubble(key = id, index = date, coords = c(long, lat))
R> cubble_nested
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts                
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_nested |>
+   mutate(rain_day = sum(ts$prcp != 0))
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts       rain_day
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>      <int>
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble>      104
2 ASN00010311 -31.9  117. 179   york            94623 <tibble>       89
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble>       90
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble>      106
5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble>       63
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_long <- cubble_nested |> face_temporal()
R> cubble_long
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin
  <chr>       <date>     <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3
2 ASN00009021 2020-01-02     0  24.9  16.4
3 ASN00009021 2020-01-03     6  23.2  13  
4 ASN00009021 2020-01-04     0  28.4  12.4
5 ASN00009021 2020-01-05     0  35.3  11.6
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_back <- cubble_long |> face_spatial()
R> cubble_back
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts                
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
\end{CodeOutput}
\begin{CodeInput}
R> identical(cubble_nested, cubble_back)
\end{CodeInput}
\begin{CodeOutput}
[1] TRUE
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_unfold <- cubble_long |> unfold(long, lat)
R> cubble_unfold
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin  long   lat
  <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3  116. -31.9
2 ASN00009021 2020-01-02     0  24.9  16.4  116. -31.9
3 ASN00009021 2020-01-03     6  23.2  13    116. -31.9
4 ASN00009021 2020-01-04     0  28.4  12.4  116. -31.9
5 ASN00009021 2020-01-05     0  35.3  11.6  116. -31.9
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_unfold |> face_spatial() |> face_temporal()
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin
  <chr>       <date>     <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3
2 ASN00009021 2020-01-02     0  24.9  16.4
3 ASN00009021 2020-01-03     6  23.2  13  
4 ASN00009021 2020-01-04     0  28.4  12.4
5 ASN00009021 2020-01-05     0  35.3  11.6
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # example with a tsibble created from climate_flat
R> raw <- climate_flat |> tsibble::as_tsibble(key = id, index = date)
R> dt <-  raw |> cubble::as_cubble(coords = c(long, lat))
R> dt
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts                
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tbl_ts [366 x 4]>
2 ASN00010311 -31.9  117. 179   york            94623 <tbl_ts [366 x 4]>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tbl_ts [366 x 4]>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tbl_ts [366 x 4]>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tbl_ts [366 x 4]>
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> covid |> head(5)
\end{CodeInput}
\begin{CodeOutput}
# A tsibble: 5 x 5 [1D]
# Key:       lga [1]
# Groups:    lga, source [1]
  date       lga        source                            n roll_mean
  <date>     <chr>      <chr>                         <int>     <dbl>
1 2022-01-01 Alpine (S) Contact with a confirmed case     1        NA
2 2022-01-02 Alpine (S) Contact with a confirmed case     2        NA
3 2022-01-03 Alpine (S) Contact with a confirmed case     4        NA
4 2022-01-04 Alpine (S) Contact with a confirmed case     4        NA
5 2022-01-05 Alpine (S) Contact with a confirmed case     2        NA
\end{CodeOutput}
\begin{CodeInput}
R> lga |> head(5)
\end{CodeInput}
\begin{CodeOutput}
Simple feature collection with 5 features and 7 fields
Geometry type: MULTIPOLYGON
Dimension:     XY
Bounding box:  xmin: 142.3535 ymin: -38.67876 xmax: 147.3909 ymax: -36.39269
Geodetic CRS:  WGS 84
    lga_code_2018            lga state_code_2016 state_name_2016 areasqkm_2018
132         20110     Alpine (S)               2        Victoria     4788.1568
133         20260    Ararat (RC)               2        Victoria     4211.1171
134         20570   Ballarat (C)               2        Victoria      739.0321
135         20660    Banyule (C)               2        Victoria       62.5402
136         20740 Bass Coast (S)               2        Victoria      865.8095
    cent_long  cent_lat                       geometry
132  146.9742 -36.85357 MULTIPOLYGON (((146.7258 -3...
133  142.8432 -37.47271 MULTIPOLYGON (((143.1807 -3...
134  143.7815 -37.49286 MULTIPOLYGON (((143.6622 -3...
135  145.0851 -37.73043 MULTIPOLYGON (((145.1357 -3...
136  145.5581 -38.50730 MULTIPOLYGON (((145.5207 -3...
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cb <- as_cubble(
+   list(spatial = lga, temporal = covid),
+   key = lga, index = date, coords = c(cent_long, cent_lat)
+   )
\end{CodeInput}
\begin{CodeOutput}
! Some sites in the temporal table don't have corresponding spatial information
\end{CodeOutput}
\begin{CodeOutput}
! Some sites in the spatial table don't have corresponding temporal information
\end{CodeOutput}
\begin{CodeOutput}
! Use argument `output = "unmatch"` to check on the unmatched key
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeOutput}
# cubble:   id [8]: nested form
# bbox:     [144.52, -37.73, 146.06, -36.55]
# temporal: date [date], matched_var [dbl]
  id          name                  lat  long type   dist group ts       n_match
  <chr>       <chr>               <dbl> <dbl> <chr> <dbl> <int> <list>     <int>
1 405234      SEVEN CREEKS @ D/S~ -36.9  146. river  6.15     5 <tibble>      21
2 404207      HOLLAND CREEK @ KE~ -36.6  146. river  8.54    10 <tibble>      21
3 ASN00082042 strathbogie         -36.8  146. clim~  6.15     5 <tibble>      21
4 ASN00082170 benalla airport     -36.6  146. clim~  8.54    10 <tibble>      21
5 230200      MARIBYRNONG RIVER ~ -37.7  145. river  6.17     6 <tibble>      19
6 ASN00086038 essendon airport    -37.7  145. clim~  6.17     6 <tibble>      19
7 406213      CAMPASPE RIVER @ R~ -37.0  145. river  1.84     1 <tibble>      18
8 ASN00088051 redesdale           -37.0  145. clim~  1.84     1 <tibble>      18
\end{CodeOutput}
\end{CodeChunk}
